---
output: 
  github_document:
    toc: true
    toc_depth: 2
  html_document:
    toc: true
    theme: united
    keep_md: true
---

# osbma
<!-- badges: start -->
[![License: GPL v2](https://img.shields.io/badge/License-GPL_v2-blue.svg)](https://www.gnu.org/licenses/old-licenses/gpl-2.0.en.html)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
[![R-CMD-check](https://github.com/sidiwang/osbma/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sidiwang/osbma/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


The aim of the **osbma** R package is to predict overall survival in solid tumor oncology studies using joint models combined through Bayesian model averaging. 

In solid tumor oncology studies, overall survival (OS) is the definitive endpoint for accessing patient benefit. Reliable OS predictions are very valuable for both drug development and patients' end-of-life medical care. Existing methods often ignore the underlying process of tumor burden measurements and incur "information loss". Motivated by an advanced renal cell carcinoma (RCC) clinical trial, in this paper, we propose a multivariate joint modeling approach to assess the underlying dynamics of the progression-free survival (PFS) components to forecast the death times of trial participants. Through Bayesian model averaging, our proposed method improves the accuracy of the OS forecast by combining joint models developed based on each granular component of PFS. A case study of the RCC trial is conducted, and our method provides the most accurate predictions across all tested scenarios. The reliability of our proposed method is also verified through extensive simulation studies, which include the scenario where OS is completely independent of PFS. Overall, the proposed methodology provides a promising candidate for reliable OS prediction in solid tumor oncology studies. 

## Installation

Install the development version from GitHub:
```{r echo = TRUE, eval = FALSE}
# Install devtools first if you haven't done so
library(devtools)
# install osbma
devtools::install_github("sidiwang/osbma")
library(osbma)
```


## **osbma** functions covered in this package

* The main function that generates posterior samples based on user input trial dataset  - `osbma`
* Overall survival prediction based on posterior samples generated through MCMC - `predict.osbma`
* Plots generated based on predicted overall survival - `plot.predict.data`

## Example 
`osbma`:
We call the `osbma` function using data from an oncology trial with 400 total individuals. 
```{r}
data <- osbma::data
result <- osbma::osbma(data = data, covariate = "trt", method = "rjags")
options(max.print = 40)
summary(result)
```
`predict.osbma`:
Here, we call the prediction function to predict the overall survival of all patients:
```{r}
prediction <- predict(result)
```
`plot.predict.osbma`:
The plot function provides 3 types of plots:
* the Kaplan-Meier curve of overall survival
```{r}
plot(prediction, trt.col.name = "trt", type = "KM")
```
* the patients' survival timeline of the trial
```{r}
# plot(prediction, trt.col.name = "trt", type = "date")
```
* the estimated time of the nth death in the trial (here we use the 120th death as an example)
```{r}
# plot(prediction, trt.col.name = "trt", type = 120)
```

As `.md` file doesn't show interactive plots, please check file `demo.Rmd` for the last two plots. Please download `demo.Rmd` and view it in browser. 

